---
title: "Statistical analysis"
author: "JH, CS"
date: "2024-02-09"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(tidy = FALSE)
options(width = 120)
```

# 1. Software versions

## 1.1. R version
```{r}
version
```

## 1.2. Package versions
```{r}
library(tidyverse)
packageVersion("tidyverse")
```

```{r}
library(readxl)
packageVersion("readxl")
```

```{r}
library(crosstable)
packageVersion("crosstable")
```

```{r}
library(reshape2)
packageVersion("reshape2")
```

```{r}
library(Hmisc)
packageVersion("Hmisc")
```

```{r}
library(nortest)
packageVersion("nortest")
```

```{r}
library(ggpubr)
packageVersion("ggpubr")
```

```{r}
library(rcompanion)
packageVersion("rcompanion")
```

# 2. Plot aesthetics
```{r include=TRUE}
pie_theme <- theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_blank(),
    panel.grid=element_blank(),
    axis.ticks = element_blank(),
    plot.title=element_text(size=14, face="bold")
  ) +
  theme(axis.text.x=element_blank())

plot_theme<- theme_minimal() +
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.line = element_line(colour="black", linewidth=.75),
    axis.ticks = element_line(colour="black", linewidth=.75),
    axis.text = element_text(size=10, face="bold"),
    axis.title = element_text(size=12, face="bold"),
    legend.text = element_text(size=10, face="bold"),
    legend.title = element_text(size=12, face="bold")
  )
```

# 3. Data import
```{r}
exceldata <- read_excel("results-survey.xlsx")                          
Data_Tabl<- data.frame(exceldata)
```

## 3.1. Data cleaning & overview

### 3.1.1. Standardization of row names and value language
```{r}
colnames(Data_Tabl) <- c(
     "lastpage",
     "Image_seen_before",
     "Age",
     "Gender",
     "F15003",
     "F03004",
     "B001",
     "B003",
     "F03002",
     "B002",
     "B004",
     "F15004",
     "F03003",
     "F03001",
     "B005",
     "F15001",
     "B006",
     "F15002",
     "B007",
     "B008",
     "interviewtime",
     "grouptime_general_questions",
     "E001_time",
     "E002_time",
     "E003_time",
     "grouptime_image_classification",
     "F15003_time",
     "F03004_time",
     "B001_time",
     "B003_time",
     "F03002_time",
     "B002_time",
     "B004_time",
     "F15004_time",
     "F03003_time",
     "F03001_time",
     "B005_time",
     "F15001_time",
     "B006_time",
     "F15002_time",
     "B007_time",
     "B008_time"
     )

# translate text values to English
Data_Tabl$Image_seen_before <- 
  dplyr::recode(Data_Tabl$Image_seen_before, ja="yes", nein="no")
Data_Tabl$Gender <- 
  dplyr::recode(Data_Tabl$Gender, männlich="male", weiblich="female")

Data_Tabl <- Data_Tabl %>%
  dplyr::mutate(across(c(F15003,F03004,F03002,F15004,F03003,F03001,F15001,F15002,B001,B003,B002,B004,B005,B006,B007,B008), \(x) dplyr::recode(x,"KI-generiert"="AI-generated", "Echt"="genuine")))

Data_Tabl$ID <- 1:nrow(Data_Tabl)
```

### 3.1.2. Remove survey test runs
```{r}
# drop survey test runs
Data_Tabl <- Data_Tabl[Data_Tabl$Age != "test",]
```

### 3.1.3. Study completion
```{r}
# participants who just opened study without answering any question
Data_Tabl$lastpage[is.na(Data_Tabl$lastpage)] <- 0

# plot overview
histCompletion <- ggplot(Data_Tabl, aes(x=lastpage)) +
  geom_histogram(binwidth=1) +
  xlab("Last answered question") +
  ylab("Participant count")

histCompletion
```

##### Remove participants who didn't complete the study
Number of participants starting the study
```{r}
n_all <- nrow(Data_Tabl)
n_all
```

Number of participants finishing the study
```{r}
Data_Tabl <- Data_Tabl[Data_Tabl$lastpage == 19,]
n_complete <- nrow(Data_Tabl)
n_complete
```
Percent completed
```{r}
p_complete <- n_complete/n_all*100
p_complete
```
## 3.2. Demographics

##### Age
```{r}
# convert age to integer
Data_Tabl$Age <- Data_Tabl$Age %>% strtoi()

# mark nonsense age entries
Data_Tabl$Age[Data_Tabl$Age>110] = NA

# plot overview
histAge <- ggplot(Data_Tabl, aes(x=Age)) +
  geom_histogram(binwidth = 1)

histAge
```


##### Gender <> Prior experience
```{R}
gender <- as.data.frame(table(addNA(Data_Tabl$Image_seen_before),addNA(Data_Tabl$Gender)))
colnames(gender) <- c("Group", "Gender", "Frequency")


ggplot(gender, aes(x = Group, y = Frequency, fill=Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Participants by Gender",
       x = "Histological image seen before",
       y = "Frequency",
       fill = "Gender") 
```

##### Gender <> Prior experience
```{r}
crosstable(Data_Tabl, c(Image_seen_before), by=c(Gender), funs=c(sum)) %>% 
  as_flextable(keep_id=TRUE)
```

##### Age <> Prior experience
```{r}
ggplot(Data_Tabl, aes(x = Age), na.rm = TRUE) +
  geom_histogram(binwidth = 5, fill = "darkblue", color="grey", na.rm = TRUE) +
  facet_wrap(~ Image_seen_before) +
  labs(title = "Participants by age having vs. not having seen histological image before",
       x = "Age",
       y = "Frequency",
       fill = "Histological image seen before") +
    scale_fill_brewer(palette="Set1")
```

##### Age <> Prior experience
```{r}
crosstable(drop_na(Data_Tabl, Age), c(Age), by=c(Image_seen_before), funs=c(median, mean, "std dev"=sd)) %>% as_flextable(keep_id=TRUE)
```

##### Prior experience
```{r}
n_participants_experience <- table(addNA(Data_Tabl$Image_seen_before))
p_participants_experience <- n_participants_experience / length(Data_Tabl$Image_seen_before) *100

expTable <- rbind(n_participants_experience, p_participants_experience)
colnames(expTable) <- c("naive", "expert", "NaN")
rownames(expTable) <- c("count", "percent")

expTable
```

## 3.3. Handling datasets containing missing values ('NA')

### 3.3.1. Missing values for classification duration
Replace duration data with NA in cases where an image has not been classified.
```{r}
Data_Tabl$B001_time[is.na(Data_Tabl$B001)]<-NA
Data_Tabl$B002_time[is.na(Data_Tabl$B002)]<-NA
Data_Tabl$B003_time[is.na(Data_Tabl$B003)]<-NA
Data_Tabl$B004_time[is.na(Data_Tabl$B004)]<-NA
Data_Tabl$B005_time[is.na(Data_Tabl$B005)]<-NA
Data_Tabl$B006_time[is.na(Data_Tabl$B006)]<-NA
Data_Tabl$B007_time[is.na(Data_Tabl$B007)]<-NA
Data_Tabl$B008_time[is.na(Data_Tabl$B008)]<-NA
Data_Tabl$F03001_time[is.na(Data_Tabl$F03001)]<-NA
Data_Tabl$F03002_time[is.na(Data_Tabl$F03002)]<-NA
Data_Tabl$F03003_time[is.na(Data_Tabl$F03003)]<-NA
Data_Tabl$F03004_time[is.na(Data_Tabl$F03004)]<-NA
Data_Tabl$F15001_time[is.na(Data_Tabl$F15001)]<-NA
Data_Tabl$F15002_time[is.na(Data_Tabl$F15002)]<-NA
Data_Tabl$F15003_time[is.na(Data_Tabl$F15003)]<-NA
Data_Tabl$F15004_time[is.na(Data_Tabl$F15004)]<-NA
```

### 3.3.2. Missing values for prior experience
Preparation of the analysis table: Assignment of lines answering NA to the question E001 ("Image_seen_before").

Rationale: Those who did not know histological images are likely to even not knowing the term and, therefore might have answered NA.
```{r}
Data_Tabl$Image_seen_before[is.na(Data_Tabl$Image_seen_before)]<-"no"

n_participants_experience_noNA <- table(addNA(Data_Tabl$Image_seen_before))
p_participants_experience_noNA <- n_participants_experience_noNA / length(Data_Tabl$Image_seen_before)

expTable_noNA <- rbind(n_participants_experience_noNA, p_participants_experience_noNA)
colnames(expTable_noNA) <- c("naive", "expert", "nan")
rownames(expTable_noNA) <- c("n", "p")

n_naive <- expTable_noNA[1,1]
n_expert <- expTable_noNA[1,2]
expTable_noNA
```

### 3.3.3. Missing values for classifications
Skipped answers during classification are considered "misclassified", i.e. 
* for artificial images 'genuine' replaces 'NA'
* for genuine images 'AI-generated' replaces 'NA'

##### Percent of non-classified images (i.e. skipped images in survey)
```{R}
na_entries_count <- sum(is.na(Data_Tabl$B001)) +
                    sum(is.na(Data_Tabl$B002)) +
                    sum(is.na(Data_Tabl$B003)) +
                    sum(is.na(Data_Tabl$B004)) +
                    sum(is.na(Data_Tabl$B005)) +
                    sum(is.na(Data_Tabl$B006)) +
                    sum(is.na(Data_Tabl$B007)) +
                    sum(is.na(Data_Tabl$B008)) +
                    sum(is.na(Data_Tabl$F03001)) +
                    sum(is.na(Data_Tabl$F03002)) +
                    sum(is.na(Data_Tabl$F03003)) +
                    sum(is.na(Data_Tabl$F03004)) +
                    sum(is.na(Data_Tabl$F15001)) +
                    sum(is.na(Data_Tabl$F15002)) +
                    sum(is.na(Data_Tabl$F15003)) +
                    sum(is.na(Data_Tabl$F15004))

entries_total <- nrow(Data_Tabl)*16
percent_na <- na_entries_count/entries_total*100
percent_na
```

##### Relabel non-classified images as misclassified
```{r}
Data_Tabl <- Data_Tabl %>%
  mutate_at(vars(F15003,F03004,F03002,F15004,F03003,F03001,F15001,F15002), ~ifelse(is.na(.), "genuine", .))
  
Data_Tabl <- Data_Tabl %>%
  mutate_at(vars(B001,B003,B002,B004,B005,B006,B007,B008), ~ifelse(is.na(.), "AI-generated", .))
```

##### Re-calculate age after treatment of missing values
```{r}
crosstable(drop_na(Data_Tabl, Age), c(Age), by=c(Image_seen_before), funs=c(median, mean, "std dev"=sd)) %>% as_flextable(keep_id=TRUE)
```

### 3.3.4. Missing values for gender
Replacing 'NA' entries in the gender column with 'unknown'
```{r}
Data_Tabl$Gender[is.na(Data_Tabl$Gender)]<-"unknown"
```

## 3.4. Format data

##### melt()
```{r}
classificationLong_response <- melt(Data_Tabl[,c(2:20, 43)], id = c("ID", "Image_seen_before", "Age", "Gender"))
classificationLong_time <- melt(Data_Tabl[,c(2:4, 27:43)], id = c("ID", "Image_seen_before", "Age", "Gender"))
imageClassificationDataLong <- cbind(classificationLong_response, classificationLong_time$value)
colnames(imageClassificationDataLong)[7] <- "time"
```

##### Add columns: generation, trainingBase, response, stimNo
```{r}
# generation: ground truth identity: artificial vs. genuine
imageClassificationDataLong$variable <- as.character(imageClassificationDataLong$variable)
imageClassificationDataLong <- imageClassificationDataLong %>%
  mutate(
    generation = case_when(
      startsWith(variable,"F") ~ "AI-generated",
      startsWith(variable,"B") ~ "genuine"
    )
  )

# trainingBase: ground truth identity: genuine (B0), trained using 3 (F3), trained using 15 F(15) genuine samples
imageClassificationDataLong <- imageClassificationDataLong %>%
  mutate(
    trainingBase = case_when(
      startsWith(variable,"F15") ~ "F15",
      startsWith(variable,"F03") ~ "F3",
      startsWith(variable,"B") ~ "B0"
    )
  )

# response: correct vs. incorrect classification
imageClassificationDataLong <- imageClassificationDataLong %>%
  mutate(
    response = case_when(
      value == generation ~ "correct",
      value != generation ~ "incorrect"
    )
  )

# response_i: same as response, but as integer
imageClassificationDataLong <- imageClassificationDataLong %>%
  mutate(
    response_i = case_when(
      value == generation ~ 1,
      value != generation ~ 0
    )
  )

# stimNo: number of stimulus in presentation during study
imageClassificationDataLong <- imageClassificationDataLong %>%
  mutate(
    stimNo = case_when(
      variable == "F15003" ~ 1,              
      variable == "F03004" ~ 2,
      variable == "B001" ~ 3,
      variable == "B003" ~4,
      variable == "F03002" ~ 5,
      variable == "B002" ~ 6,
      variable == "B004" ~ 7,
      variable == "F15004" ~ 8,
      variable == "F03003" ~ 9,
      variable == "F03001" ~ 10,
      variable == "B005" ~ 11,
      variable == "F15001" ~ 12,
      variable == "B006" ~ 13,
      variable == "F15002" ~ 14,
      variable == "B007" ~ 15,
      variable == "B008" ~ 16
    )
  )

```

# 4. Analysis of classification performance

## 4.1.  Analysis of the distinguishability of artificial vs. genuine images
Guiding question: Can the real images be well distinguished from the generated ones?

### 4.1.1. Format data
```{r}
tbl_a = table(imageClassificationDataLong[,c("generation","value")])
tbl_a <- addmargins(tbl_a)
tbl_a
```

### 4.1.2. F1 Score
```{r}
TP <- tbl_a["AI-generated", "AI-generated"]
FN <- tbl_a["AI-generated", "genuine"]
FP <- tbl_a["genuine", "AI-generated"]
TN <- tbl_a["genuine", "genuine"]

precision <- TP/(TP+FP)
sens <- TP/(TP+FN)

F1_score <- 2/(1/precision + 1/sens) 
F1_score
```

## 4.2. Confidence intervals

### 4.2.1. Naive participants
```{r}
# gather data for naives
dat_unexperienced <- filter(imageClassificationDataLong,Image_seen_before=="no")

tbl_SensSpec_n = table(dat_unexperienced[,c("generation","value")])
tbl_SensSpec_n <- addmargins(tbl_SensSpec_n)

TP_n <- tbl_SensSpec_n["AI-generated", "AI-generated"]
FN_n <- tbl_SensSpec_n["AI-generated", "genuine"]
FP_n <- tbl_SensSpec_n["genuine", "AI-generated"]
TN_n <- tbl_SensSpec_n["genuine", "genuine"]


# Sensitivity + specificity: Confidence intervals  using the Wilson Score Interval.
allpos <- TP_n + FN_n
allneg <- TN_n + FP_n

sensitivity_ci_n <- binconf(TP_n,allpos,method="wilson")
specificity_ci_n <- binconf(TN_n,allneg,method="wilson")
```

```{r}
tbl_SensSpec_n
```

```{r}
print("Sensitivity Confidence Interval:")
print(sensitivity_ci_n)
```

```{r}
print("Specificity Confidence Interval:")
print(specificity_ci_n)
```

### 4.2.2. Expert participants
```{r}
# gather data for expert participants
dat_unexperienced <- filter(imageClassificationDataLong,Image_seen_before=="yes")

tbl_SensSpec_e = table(dat_unexperienced[,c("generation","value")])
tbl_SensSpec_e <- addmargins(tbl_SensSpec_e)

TP_e <- tbl_SensSpec_e["AI-generated", "AI-generated"]
FN_e <- tbl_SensSpec_e["AI-generated", "genuine"]
FP_e <- tbl_SensSpec_e["genuine", "AI-generated"]
TN_e <- tbl_SensSpec_e["genuine", "genuine"]


# Sensitivity + specificity: Confidence intervals  using the Wilson Score Interval.
allpos_e <- TP_e + FN_e
allneg_e <- TN_e + FP_e

sensitivity_ci_e <- binconf(TP_e,allpos_e,method="wilson")
specificity_ci_e <- binconf(TN_e,allneg_e,method="wilson")
```

```{r}
tbl_SensSpec_e
```

```{r}
print("Sensitivity Confidence Interval:")
print(sensitivity_ci_e)
```

```{r}
print("Specificity Confidence Interval:")
print(specificity_ci_e)
```

### 4.2.3. Plot results
```{r}
df_sensSpec <- data.frame(
  subject = c("naive", 
              "naive", 
              "expert", 
              "expert"),
  measure = c("Sens", 
              "Spec", 
              "Sens", 
              "Spec"),
  estimate = c(sensitivity_ci_n[1], 
               specificity_ci_n[1], 
               sensitivity_ci_e[1], 
               specificity_ci_e[1]),
  CI_lower = c(sensitivity_ci_n[2], 
               specificity_ci_n[2], 
               sensitivity_ci_e[2], 
               specificity_ci_e[2]),
  CI_upper = c(sensitivity_ci_n[3], 
               specificity_ci_n[3], 
               sensitivity_ci_e[3], 
               specificity_ci_e[3])
  )

ggplot(df_sensSpec, aes(x=subject, y=estimate, colour=measure, fill=measure)) +
  geom_point(position=position_dodge(0.185), size=5, shape=16) +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), width=0, size=1, position=position_dodge(0.185)) +
  scale_color_manual(values =  c("#16d108", "#0756de")) +
  labs(y= "Respones [%]", x = "") +
  plot_theme
```

## 4.3. Overall performance 

### 4.3.1. Naive participants

##### Numbers
```{r}
n_correct <- TP_n+TN_n
n_correct

n_incorrect <- FP_n+FN_n
n_incorrect

p_correct <- (TP_n+TN_n) / (TP_n+TN_n+FP_n+FN_n) *100
p_correct

p_incorrect <- (FP_n+FN_n) / (TP_n+TN_n+FP_n+FN_n) *100
p_incorrect
```

##### Pie chart proportions
```{r}
# pie chart
df_n_all <- data.frame(
  group = c("Correct", "False"),
  value = c(TP_n+TN_n, FP_n+FN_n)
)

df_n_all$value <- df_n_all$value / sum(df_n_all$value) * 100

bp_n_all <- ggplot(df_n_all, aes(x="", y=value, fill=group)) +
  geom_bar(width=1, stat="identity") +
  coord_polar("y", start=0) + 
  pie_theme +
  scale_fill_manual(values = c("#03c7ff", "#fc7303"))

bp_n_all

```

##### Statistics vs. chance
Binomial exact test: Performance significantly better than chance
```{r}
binom_n <- binom.test(TP_n+TN_n, TP_n+TN_n+FP_n+FN_n, p = 0.5,
                       alternative = c("two.sided", "less", "greater"),
                       conf.level = 0.95)

binom_n
```

##### Performance genuine vs. artificial
Performance naive - genuine:
```{r}
n_correct <- TN_n
n_correct

n_incorrect <- FP_n
n_incorrect

p_correct <- (TN_n) / (TN_n+FP_n) *100
p_correct

p_incorrect <- (FP_n) / (FP_n+TN_n) *100
p_incorrect
```

Performance naive - artificial:
```{r}
n_correct <- TP_n
n_correct

n_incorrect <- FN_n
n_incorrect

p_correct <- (TP_n) / (TP_n+FN_n) *100
p_correct

p_incorrect <- (FN_n) / (TP_n+FN_n) *100
p_incorrect
```

Chi2 test: No statistical difference for performance of naive participants between image categories.
```{r}
correct_n <- c(TP_n, TN_n)
incorrect_n <- c(FP_n, FN_n)

cnttab_n <- matrix(c(correct_n, incorrect_n), ncol=2)
rownames(cnttab_n) <- c("Artificial", "Genuine")
colnames(cnttab_n) <- c("Correct", "False")
chi2_n <- chisq.test(cnttab_n)
print(cnttab_n)
print(chi2_n)
```

### 4.3.2. Expert participants

##### Numbers
```{r}
n_correct <- TP_e+TN_e
n_correct

n_incorrect <- FP_e+FN_e
n_incorrect

p_correct <- (TP_e+TN_e) / (TP_e+TN_e+FP_e+FN_e) *100
p_correct

p_incorrect <- (FP_e+FN_e) / (TP_e+TN_e+FP_e+FN_e) *100
p_incorrect
```

##### Pie chart proportions
```{r}
df_e_all <- data.frame(
  group = c("Correct", "False"),
  value = c(TP_e+TN_e, FP_e+FN_e)
)

df_e_all$value <- df_e_all$value / sum(df_e_all$value) * 100

bp_e_all <- ggplot(df_e_all, aes(x="", y=value, fill=group)) +
  geom_bar(width=1, stat="identity") +
  coord_polar("y", start=0) + 
  pie_theme +
  scale_fill_manual(values = c("#03c7ff", "#fc7303"))

bp_e_all
```

##### Statistics vs. naive participants
Experts perform statistically significantly better than naive participants.
```{r}
exprcd <- c((TP_e+TN_e),(FN_e+FP_e))
unexprcd <- c((TP_n+TN_n),(FN_n+FP_n))

cnttab <- matrix(c(exprcd,unexprcd), nrow = 2)
rownames(cnttab) <- c("Experienced", "Unexperienced")
colnames(cnttab) <- c("Correct", "Incorrect")

chi_quadrat_test <- chisq.test(cnttab)

print(cnttab)
print(chi_quadrat_test)
```
##### Performance genuine vs. artificial
Performance expert - genuine:
```{r}
n_correct <- TN_e
n_correct

n_incorrect <- FP_e
n_incorrect

p_correct <- (TN_e) / (TN_e+FP_e) *100
p_correct

p_incorrect <- (FP_e) / (FP_e+TN_e) *100
p_incorrect
```

Performance expert - artificial:
```{r}
n_correct <- TP_e
n_correct

n_incorrect <- FN_e
n_incorrect

p_correct <- (TP_e) / (TP_e+FN_e) *100
p_correct

p_incorrect <- (FN_e) / (TP_e+FN_e) *100
p_incorrect
```
Chi2 test: No statistical difference for performance of expert participants between image categories.
```{r}
correct_e <- c(TP_e, TN_e)
incorrect_e <- c(FP_e, FN_e)

cnttab_e <- matrix(c(correct_e, incorrect_e), ncol=2)
rownames(cnttab_e) <- c("Artificial", "Genuine")
colnames(cnttab_e) <- c("Correct", "False")
chi2_e <- chisq.test(cnttab_e)
print(cnttab_e)
print(chi2_e)
```

## 4.4. Performance by image category

### 4.4.1. Bar plot proportions
```{r}
imageDataCast <- dcast(imageClassificationDataLong, trainingBase + Image_seen_before ~ response, length)
imageDataCast$pCorrect <- imageDataCast$correct / (imageDataCast$correct + imageDataCast$incorrect)

bp_sup <- ggplot(imageDataCast, aes(x=Image_seen_before, y=pCorrect*100, fill=fct_reorder(trainingBase, c(5,1,3,6,2,4)))) +
  geom_bar(width=0.9, stat="identity", position="dodge") +
    scale_fill_manual(values = c("#0c1ff0", "#00ff00", "#0de609")) +
  labs(y= "Correct respones [%]", x = "") +
  ylim(0, 100) + 
  plot_theme

bp_sup
```

### 4.4.2. Statistics: Binomial regression

##### Model
```{r}
model <- glm(factor(response_i) ~ Image_seen_before * trainingBase,
            data=imageClassificationDataLong,
            family="binomial")
summary(model)
```

##### Estimate + confidence intervals
```{r}
as.data.frame(cbind(coef(model), confint(model)))
```

##### Odds ratio + confidence intervals
```{r}
OR_vals <- as.data.frame(exp(cbind(coef(model), confint(model))))
OR_vals
```

##### Plot results
```{r}
OR_vals$no <- c(1:6)
colnames(OR_vals) <- c("pnt", "lower", "upper", "no")

ggplot(OR_vals, aes(y=pnt, x=factor(no))) +
  geom_point(position=position_dodge(0.185), size=5, shape=16) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width=0, size=1, position=position_dodge(0.185)) +
  labs(y= "OR + 95% CI", x = "") +
  scale_x_discrete(labels=c(rownames(OR_vals))) +
  plot_theme +
  theme(axis.text.x = element_text(angle=90)) +
  geom_hline(yintercept=1)
```

## 4.5. Participant-wise performance

##### Remove images without response
```{r}
nrow(imageClassificationDataLong)
imageClassificationDataLong_2 <- imageClassificationDataLong %>% drop_na(time)
nrow(imageClassificationDataLong_2)
```

### 4.5.1. Format data
```{r}
# preallocate
paticipant_data_length <- length(unique(imageClassificationDataLong_2$ID))
participant_data <- data.frame(no=rep(NA,paticipant_data_length),
                               id=rep(NA,paticipant_data_length),
                               experience=rep(NA,paticipant_data_length),
                               mean_response=rep(NA,paticipant_data_length))


counter = 1
for(x in unique(imageClassificationDataLong_2$ID)){
      tmp_data <- imageClassificationDataLong_2 %>% subset(ID == x)
      participant_data$no[counter] <- counter
      participant_data$id[counter] <- tmp_data$ID[1]
      participant_data$experience[counter] <- tmp_data$Image_seen_before[1]
      participant_data$mean_response[counter] <- mean(tmp_data$response_i)
      counter = counter + 1
}

```

### 4.5.2. Plot results as cumulative plot
```{r}
ggplot(participant_data, aes(linetype=experience, color=experience, x=mean_response)) +
  stat_ecdf(size=0.75) +
  plot_theme +
  xlab("Participant-wise performance [% correct]") +
  scale_linetype_manual(values=c("longdash", "solid"))+
  scale_color_manual(values=c("black", "#858382")) +
  ylab("CDF") +
  ylim(-0.01, 1.01) + xlim(-0.01, 1.01)

```

### 4.5.3. Plot results as boxplot
```{r}
ggplot(participant_data, aes(fill=experience, y=mean_response)) +
  geom_boxplot(notch = TRUE) +
  plot_theme +
  xlab("Participant-wise performance [% correct]") +
  scale_fill_manual(values=c("black", "#858382")) +
  ylab("Performance") + 
  ylim(-0.01, 1.01)
```

### 4.5.4. Statistical comparison
##### Anderson-Darling test for normality
```{r}
ad.test(participant_data$mean_response)
ad.test(participant_data$mean_response[participant_data$experience=="yes"])
ad.test(participant_data$mean_response[participant_data$experience=="no"])
```
RESULTS: The distribution of the data differs highly significantly from a normal distribution.

##### Inferential statistics: Wilcoxon rank sum test
```{r}
med_perf_e <- median(participant_data$mean_response[participant_data$experience == "yes"])
med_perf_n <- median(participant_data$mean_response[participant_data$experience == "no"])
c(med_perf_e, med_perf_n)
                    

wilcox.test(participant_data$mean_response[participant_data$experience == "yes"],
            participant_data$mean_response[participant_data$experience == "no"], paired=FALSE)
```
RESULTS: Participant-wise performance differs statistically significantly between naive and expert participants.

# 5. Analysis of response times
## 5.1. Remove images without response
as there is no associated response time to work with
```{r}
nrow(imageClassificationDataLong)
imageClassificationDataLong_2 <- imageClassificationDataLong %>% drop_na(time)
nrow(imageClassificationDataLong_2)
```

## 5.2. Sanity checks
### 5.2.1. Check effect of presentation order on response time
```{r}
bp_order_time <- ggplot(imageClassificationDataLong_2, aes(y=time, x=as.factor(stimNo), fill=generation)) +
  geom_violin() +
  stat_summary(fun="median") +
  ylim(0,30) +
  xlab("Stimulus number during presentation") +
  ylab("Response time distribution and median [s]")

bp_order_time
```

INTERPRETATION: There is a clear, effect of stimulus presentation order on response times. This results in a higher response time in particular for the first two images presented, which are both artificial images. Thus do not include image category as an factor for analysis of response times.

NOTE:  The 371 dropped values in this in subsequent plots are those response times > 30s, which do not get plotted due to the imposed axis limit:
```{r}
sum(imageClassificationDataLong_2$time>30)
```

However, these 'outliers' are still included in all statistical analyses. 

### 5.2.2. Effect of presentation order on classification correctness

##### Overview
```{r}
bp_order_response <- ggplot(imageClassificationDataLong_2, aes(y=response_i, x=as.factor(stimNo), color=generation)) +
  stat_summary(fun.data = "mean_cl_boot") +
  xlab("Stimulus number during presentation") +
  ylab("Response correctness mean + CI [s]")

bp_order_response
```

##### Correlation
```{r}
dataCastByStimNo <- dcast(imageClassificationDataLong_2, stimNo ~ response, length)

dataCastByStimNo$pCorrect <- dataCastByStimNo$correct / (dataCastByStimNo$correct + dataCastByStimNo$incorrect)

ggscatter(dataCastByStimNo, x = "stimNo", y = "pCorrect",
          add = "reg.line", conf.int = TRUE,
          add.params = list(color = "blue", fill = "lightgray"))+
  stat_cor(method = "pearson")

# Note: if calculated separately by participant group the conclusion holds true as well:
# Naives: R = 0.28, p = 0.29
# Experts: R = 0.21, p = 0.43
# dataCastByStimNo <- dcast(imageClassificationDataLong_2, 
#                           stimNo ~ response + Image_seen_before, 
#                           length)
```

INTERPRETATION: There is no relevant effect of stimulus presentation order on classification correctness.

## 5.3. Statistical analysis of effects on response times

### 5.3.1. Anderson-Darling test for normality
```{r}
ad.test(imageClassificationDataLong_2$time)
```

The data are non-normally distributed. The same is true when testing for distribution of subpopulations (see below.) Thus use the Scheirer-Ray-Hare test as non-parametric alternative to an ANOVA.

### 5.3.2. Further considerations
* Age is different between experienced and controls, thus not independent of prior experience
* Gender composition is different between experienced and controls, thus not independent of prior experience
* The effect of image category depends on the stimulus presentation order

CONCLUSION: Only analyse effect of prior experience and response correctness on response time.

### 5.3.3. Scheirer-Ray-Hare test
as non-parametric alternative to ANOVA
```{r}
lm <- scheirerRayHare(time ~ Image_seen_before*response, data=imageClassificationDataLong_2)
```

```{r}
lm
```

```{r}
summary(lm)
```

### 5.3.4. Pair-wise post hoc comparisons

#### 5.3.4.1. Effect of classification correctness

##### Plot CDF: Full response times
```{r}
ggplot(imageClassificationDataLong_2, aes(color=response, x=time)) +
  stat_ecdf(size=.75) +
  xlab("Response Time [s]") +
  ylab("CDF") +
  scale_color_manual(values = c("#03c7ff", "#fc7303")) +
  plot_theme
```

INTERPRETATION: Very long tail - very likely cases in which participants just left browser window open, did something else and returned to study at at some later point.

##### Plot CDF: First 30 seconds
```{r}
ggplot(imageClassificationDataLong_2, aes(color=response, x=time)) +
  stat_ecdf(size=.75) +
  xlim(0,30) +
  xlab("Response Time [s]") +
  ylab("CDF") +
  scale_color_manual(values = c("#03c7ff", "#fc7303")) +
  plot_theme
```

##### Anderson-Darling test for normality
```{r}
ad.test(imageClassificationDataLong_2$time[imageClassificationDataLong_2$response=="correct"])
ad.test(imageClassificationDataLong_2$time[imageClassificationDataLong_2$response=="incorrect"])
```
RESULTS: The distribution of the data differs significantly from a normal distribution.

##### Wilcoxon rank sum test
```{r}
wilcox.test(imageClassificationDataLong_2$time[imageClassificationDataLong_2$response=="correct"],
       imageClassificationDataLong_2$time[imageClassificationDataLong_2$response=="incorrect"], paired=FALSE)
```

##### Median response time difference
```{r}
m_c <- median(imageClassificationDataLong_2$time[imageClassificationDataLong_2$response=="correct"])
m_i <- median(imageClassificationDataLong_2$time[imageClassificationDataLong_2$response=="incorrect"])
c(m_c, m_i)
```

#### 5.3.4.2. Effect of prior experience

##### Plot CDF: Full response times
```{r}
ggplot(imageClassificationDataLong_2, aes(linetype=Image_seen_before, x=time)) +
  stat_ecdf(size=.75) +
  xlab("Response Time [s]") +
  ylab("CDF") +
  scale_linetype_manual(values=c("longdash", "solid"))+
  plot_theme
```

##### Plot CDF: First 30 seconds
```{r}
ggplot(imageClassificationDataLong_2, aes(linetype=Image_seen_before, x=time, color=Image_seen_before)) +
  stat_ecdf(size=.75) +
  xlim(0,30) +
  xlab("Response Time [s]") +
  ylab("CDF") +
  scale_linetype_manual(values=c("longdash", "solid"))+
  scale_color_manual(values=c("black", "#858382")) +
  plot_theme
```

##### Anderson-Darling test for normality
```{r}
ad.test(imageClassificationDataLong_2$time[imageClassificationDataLong_2$Image_seen_before=="yes"])
ad.test(imageClassificationDataLong_2$time[imageClassificationDataLong_2$Image_seen_before=="no"])
```

RESULTS: The distribution of the data differs significantly from a normal distribution.

##### Wilcoxon rank sum test
```{r}
wilcox.test(imageClassificationDataLong_2$time[imageClassificationDataLong_2$Image_seen_before=="yes"],
            imageClassificationDataLong_2$time[imageClassificationDataLong_2$Image_seen_before=="no"], paired=FALSE)
```

##### Median response time difference
```{r}
m_e <-median (imageClassificationDataLong_2$time[imageClassificationDataLong_2$Image_seen_before=="yes"])
m_n <- median(imageClassificationDataLong_2$time[imageClassificationDataLong_2$Image_seen_before=="no"])
c(m_n, m_e)
```

#### 5.3.4.3. Combined effect: Prior experience <> classification correctness

##### Format data
```{r}
imageClassificationDataLong_2 <- imageClassificationDataLong_2 %>%
  mutate(
    plotGroup = case_when(
      response == "correct" & Image_seen_before == "no" ~ "naive, correct answer",
      response == "incorrect" & Image_seen_before == "no" ~ "naive, incorrect answer",
      response == "correct" & Image_seen_before == "yes" ~ "expert, correct answer",
      response == "incorrect" & Image_seen_before == "yes" ~ "expert, incorrect answer",
    )
  )
```

##### Plot CDF: First 30 seconds
```{r}
ggplot(imageClassificationDataLong_2, aes(color=plotGroup, x=time)) +
  stat_ecdf(size=.75, aes(linetype = plotGroup)) +
  xlim(0,30)+
  xlab("Response Time [s]") +
  ylab("CDF") +
  scale_color_manual(values = c("#02e1fa", "#fca503","#0690c2", "#ff8800")) +
  scale_linetype_manual(values=c("solid", "solid", "longdash", "longdash"))+
  plot_theme
```

##### Wilcoxon rank sum test with Bonferroni correction
```{r}
# data
group1 <- imageClassificationDataLong_2$time[imageClassificationDataLong_2$plotGroup == "naive, correct answer"]
group2 <- imageClassificationDataLong_2$time[imageClassificationDataLong_2$plotGroup == "naive, incorrect answer"]
group3 <- imageClassificationDataLong_2$time[imageClassificationDataLong_2$plotGroup == "expert, correct answer"]
group4 <- imageClassificationDataLong_2$time[imageClassificationDataLong_2$plotGroup == "expert, incorrect answer"]

# un-paired t-test
p1 <- wilcox.test(group1, group2, paired=FALSE)
p2 <- wilcox.test(group1, group3, paired=FALSE)
p3 <- wilcox.test(group1, group4, paired=FALSE)
p4 <- wilcox.test(group2, group3, paired=FALSE)
p5 <- wilcox.test(group2, group4, paired=FALSE)
p6 <- wilcox.test(group3, group4, paired=FALSE)

# bonferroni
p_vals <- c(p1$p.value, p2$p.value, p3$p.value, p4$p.value, p5$p.value, p6$p.value)
p_adj <- p_vals * length(p_vals)
p_adj[p_adj > 1] <- 1
p_adj
```

##### Median response time difference
```{r}
# data
m1 <- median(imageClassificationDataLong_2$time[imageClassificationDataLong_2$plotGroup == "naive, correct answer"])
m2 <- median(imageClassificationDataLong_2$time[imageClassificationDataLong_2$plotGroup == "naive, incorrect answer"])
m3 <- median(imageClassificationDataLong_2$time[imageClassificationDataLong_2$plotGroup == "expert, correct answer"])
m4 <- median(imageClassificationDataLong_2$time[imageClassificationDataLong_2$plotGroup == "expert, incorrect answer"])

median_times <- c(m1, m2, m3, m4)
median_times
```

#### 5.3.4.4. Within-participant pair-wise response time comparison

##### Format data
Within-participant pair wise comparison of average response times for correct and incorrect classifications in order to control for the possibility that people who are less well able to distinguish between image categories are slower.
```{r}
# preallocate
pairwise_data_length <- length(unique(imageClassificationDataLong_2$ID)) *
                     length(unique(imageClassificationDataLong_2$response))
pairwise_data <- data.frame(no=rep(NA,pairwise_data_length),
                          id=rep(NA,pairwise_data_length),
                          experience=rep(NA,pairwise_data_length),
                          response=rep(NA,pairwise_data_length),
                          time_mean=rep(NA,pairwise_data_length))


counter = 1
for(x in unique(imageClassificationDataLong_2$ID)){
    for(y in unique(imageClassificationDataLong_2$response)){
      tmp_data <- imageClassificationDataLong_2 %>% subset(ID == x & response == y)
      pairwise_data$no[counter] <- counter
      pairwise_data$id[counter] <- tmp_data$ID[1]
      pairwise_data$experience[counter] <- tmp_data$Image_seen_before[1]
      pairwise_data$response[counter] <- tmp_data$response[1]
      pairwise_data$time_mean[counter] <- tmp_data$time %>% mean
      counter = counter + 1
  }
}
```

Find non-paired mean values
```{r}
# missing mean times
na_ind <- which(is.na(pairwise_data$time_mean))

# add odd inds for even ones & vice versa
na_ind_addOdd <- na_ind[which(na_ind %% 2 == 0)] - 1
na_ind_addEven <- na_ind[which(na_ind %% 2 != 0)] + 1
na_ind_all <- c(na_ind, na_ind_addOdd, na_ind_addEven)
na_ind_all <- sort(setdiff(na_ind_all, na_ind_all[duplicated(na_ind_all)]))   # unique IDs after exclusion

# non-paired vs. paired data
nonpaired_data <- pairwise_data[na_ind_all,] # correct answers only
paired_data <- pairwise_data[-na_ind_all,]   # correct and incorrect answers
```

Participants excluded from pairwise comparison due to all responses correct
```{r}
# entry #s
na_ind_all
# n participant (2 entries / participant)
length(na_ind_all)/2

# by experience
tbl_all_corrrect <- table(nonpaired_data$experience[nonpaired_data$response=="correct"])
tbl_all_corrrect
```
NOTE: There are no naive participants who classify all images correctly.

##### Excursion: Enrichment statistics of experts classifying all images correctly (exact binomial test)
```{r}
# average chance of success for experts
p_expert <- mean(imageClassificationDataLong_2$response_i[imageClassificationDataLong_2$Image_seen_before=="yes"])

# probability to get all 16 images correct assuming average expert success rate
prob_all_correct <- p_expert^16

# number of experts with 16 (=100%) successes
e_all_correct <- tbl_all_corrrect[1]

# binomial test
binom.test(e_all_correct,
           n_expert,
           p=prob_all_correct)

# enrichment ratio
ExpectVal <- prob_all_correct * n_expert
ER <- 10/ExpectVal
ER
```
A total of 10 (expert) participants answered all questions correctly. This performance is way above chance.

##### Plot CDF: First 30 seconds
```{r}
ggplot(paired_data, aes(x=time_mean, color=response)) +
  stat_ecdf(size=0.75)  +
  xlab("Response Time [s]") +
  ylab("CDF") +
  scale_color_manual(values = c("#03c7ff", "#fc7303")) +
  plot_theme +
  xlim(0,30)
```

##### Plot boxplot: First 30 seconds
```{r}
ggplot(paired_data, aes(y=time_mean, fill=response)) +
  geom_boxplot(notch=TRUE)  +
  ylab("Response Time [s]") +
  scale_fill_manual(values = c("#03c7ff", "#fc7303")) +
  plot_theme +
  ylim(0,30)
```

##### Anderson-Darling test for normality
```{r}
ad.test(paired_data$time_mean)
ad.test(paired_data$time_mean[paired_data$response == "correct"])
ad.test(paired_data$time_mean[paired_data$response == "incorrect"])

```

##### Wilcoxon singed rank test
```{r}
wilcox.test(paired_data$time_mean[paired_data$response == "correct"],
            paired_data$time_mean[paired_data$response == "incorrect"], paired=TRUE)
```

##### Median response time difference
```{r}
m_c_p <- median(paired_data$time_mean[paired_data$response == "correct"])
m_i_p <- median(paired_data$time_mean[paired_data$response == "incorrect"])
c(m_c_p, m_i_p)
```
